Untitled

### 2017-02-14 Code/plot_age_standardardized_mortality_trends.Rmd

```{r load_libraries, echo=FALSE, warning=F}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)
library(plotly)
```

```{r, echo=F}
mortality.rates <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates.csv")
```

**Age-adjusted injury-related mortality (per 100,000) in white males, United States, 1969-2013**

```{r, echo=F, fig.width=11, fig.height=9}
ggplotly(ggplot(subset(mortality.rates, sex == "Male" & race == "White" & post.samp == 1 & COD == "Injuries"), aes(y=rate.per.100k, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_line(aes(col = state, lty = race)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Age-standardized mortality rate (per 100,000)") +
           xlab("Year") +
           geom_text(data = subset(mortality.rates, year == 2013 & sex == "Male" & race == "White" & post.samp == 1 & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))

ggplotly(ggplot(subset(mortality.rates, sex == "Male" & race == "White" & COD == "Injuries"), aes(y=rate.per.100k, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_line(aes(col = state, lty = factor(post.samp))) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Age-standardized mortality rate (per 100,000)") +
           xlab("Year") +
           geom_text(data = subset(mortality.rates, year == 2013 & sex == "Male" & race == "White" & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))
```

**Age-adjusted injury-related mortality (per 100,000) in black males, United States, 1969-2013**

```{r, echo=F, fig.width=11, fig.height=9}
ggplotly(ggplot(subset(mortality.rates, sex == "Male" & race == "Black" & post.samp == 1 & COD == "Injuries"), aes(y=rate.per.100k, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_line(aes(col = state, lty = race)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Age-standardized mortality rate (per 100,000)") +
           xlab("Year") +
           geom_text(data = subset(mortality.rates, year == 2013 & sex == "Male" & race == "Black" & post.samp == 1 & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))

ggplotly(ggplot(subset(mortality.rates, sex == "Male" & race == "Black" & COD == "Injuries"), aes(y=rate.per.100k, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_line(aes(col = state, lty = factor(post.samp))) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Age-standardized mortality rate (per 100,000)") +
           xlab("Year") +
           geom_text(data = subset(mortality.rates, year == 2013 & sex == "Male" & race == "Black" & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))
```

**Age-adjusted injury-related mortality (per 100,000) in white females, United States, 1969-2013**

```{r, echo=F, fig.width=11, fig.height=9}
ggplotly(ggplot(subset(mortality.rates, sex == "Female" & race == "White" & post.samp == 1 & COD == "Injuries"), aes(y=rate.per.100k, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_line(aes(col = state, lty = race)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Age-standardized mortality rate (per 100,000)") +
           xlab("Year") +
           geom_text(data = subset(mortality.rates, year == 2013 & sex == "Female" & race == "White" & post.samp == 1 & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))
```

**Age-adjusted injury-related mortality (per 100,000) in black females, United States, 1969-2013**

```{r, echo=F, fig.width=11, fig.height=9}
ggplotly(ggplot(subset(mortality.rates, sex == "Female" & race == "Black" & post.samp == 1 & COD == "Injuries"), aes(y=rate.per.100k, x = year)) + 
           geom_vline(aes(xintercept = 1983), col = "#636363", lwd = 0.5) +
           geom_vline(aes(xintercept = 1993), col = "#636363", lwd = 0.5) +
           geom_line(aes(col = state, lty = race)) +
           facet_wrap( ~ Census_Division, ncol = 3) +
           ylab("Age-standardized mortality rate (per 100,000)") +
           xlab("Year") +
           geom_text(data = subset(mortality.rates, year == 2013 & sex == "Female" & race == "Black" & post.samp == 1 & COD == "Injuries"), 
                     aes(label = stabbrs), check_overlap = T, size = 2.5))
```

>>>>

```{r load_libraries, echo=FALSE, warning=F, message=F}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)
library(plotly)
```

```{r, echo=F}
mortality.rates <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates.csv")
mortality.rates.wide <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/mortality_rates_wide.csv")
```

```{r, echo=F, fig.width=11}
for(sex_i in c("Male", "Female")) {
  for(cod in levels(mortality.rates$COD)) {
    for(region in levels(mortality.rates$Census_Region)) {
      data.sub <- subset(mortality.rates, sex == sex_i & post.samp == 1 & COD == cod  & Census_Region == region)
      a <- ggplot(data.sub, 
                      aes(y=rate.per.100k, x = year)) + 
                 geom_vline(aes(xintercept = 1983), col = "#636363", lty = 3) +
                 geom_vline(aes(xintercept = 1993), col = "#636363", lty = 3) +
                 geom_line(aes(col = state, lty = race)) +
                 facet_wrap( ~ state, ncol = 5) +
                 ylab("Age-standardized mortality rate (per 100,000)") +
                 xlab("Year") + 
                 ggtitle(paste0("Age-adjusted mortality (per 100,000) in ", sex_i, " for ", cod, " (Census Region: ", region, ")"))
      print(a)
    }
    for(region in levels(mortality.rates$Census_Region)) {    
          b <- ggplot(subset(mortality.rates.wide, sex == sex_i & post.samp == 1 & COD == cod & Census_Region == region), 
                      aes(y=rate.difference, x = year)) + 
                 geom_vline(aes(xintercept = 1983), col = "#636363", lty = 3) +
                 geom_vline(aes(xintercept = 1993), col = "#636363", lty = 3) +
                 geom_hline(aes(yintercept = 0), col = "#636363") +
                 geom_line(aes(col = state)) +
                 facet_wrap( ~ state, ncol = 5) +
                 ylab("Black-white difference (per 100,000)") +
                 xlab("Year") + 
                 ggtitle(paste0("Black-white difference in age-adjusted mortality (per 100,000) in ", 
                                sex_i, " for ", cod, " (Census Region: ", region, ")"))
          print(b)
    }
  }
}
```
